\section{Experiment setup}

\subsection{System setup}
We test the frameworks on the CentOS 7.2 server with 4 octa-core Xeon-E5 cpus and 4 GTX TITAN X(GM200) gpus.
We use Cuda 7.5 and CuDNN R5 which is the latest stable release of Cuda and CuDNN.
All deep learning frameworks are updated to latest stable release on June 2016.
The versions of the frameworks fully support CuDNN R5.
Only Torch supports the latest version of Cuda-convnet3.
The detailed system environments are represented on Table \ref{} and \ref{}.

CentOS 7.2.1511 / Linux 3.10.0-327 / Inte Xeon E5-2650@2.0GHz / 128GB DDR3@1600MHz / CuDNN 5005 / Cuda 7.5 / Torch 7 ccn2.torch, cudnn.torch R5 / Theano 0.8.2 / Caffe * / Tensorflow *
%TODO

\subsection{AlexNet model}
AlexNet \cite{} is one of the earliest successful deep neural networks on image recognition task using ImageNet dataset \cite{}.
AlexNet uses 5 convolution layers to extract features and 3 fully connected layers for classification.
Each layer has rectified linear unit(ReLU) layer for nonlinear activation.
AlexNet has been frequently used for benchmarking performance of machine learning libraries, because it utilizes most of the current DNN components such as convolution, max-pooling and dropout \cite{}.
The original AlexNet model includes Local response normalization(LRN) layer, but we exclude it for benchmarking task since LRN is very rarely used in current convolutional neural networks.
The detailed layer structure of AlexNet model on this study is presented in Table \ref{}.


\begin{table*}[]
\centering
\caption{Alexnet model used on benchmarking}
\label{alex_model}
\begin{tabular}{llllllll}
Name    & Kernel(R) & Input Channels(C) & Ouput Channels(K) & Stride(K) & Sample width(W) & Params & Flop \\
Input   &           & 3                 &                   &           & 227 x 227       &        &      \\
Conv1   & 11 x 11   & 3                 & 96                & 4         & 55 x 55         & 35K    & 55G  \\
Pool    & 3 x 3     &                   &                   & 2         &                 &        &      \\
Conv2   & 5 x 5     & 96                & 256               & 1         & 27 x 27         & 614K   & 227G \\
Pool    & 3 x 3     &                   &                   & 2         &                 &        &      \\
Conv3   & 3 x 3     & 256               & 384               & 1         & 13 x 13         & 885K   & 65G  \\
Conv4   & 3 x 3     & 384               & 384               & 1         & 13 x 13         & 1.3M   & 98G  \\
Conv5   & 3 x 3     & 384               & 256               & 1         & 13 x 13         & 885K   & 65G  \\
Pool    & 3 x 3     &                   &                   & 2         &                 &        &      \\
FC6     &           & 256 x 6 x 6       & 4096              &           &                 & 37M    & 74M  \\
FC7     &           & 4096              & 4096              &           &                 & 16M    & 32M  \\
FC8     &           & 4096              & 1000              &           &                 & 4M     & 8M   \\
Softmax &           & 1000              & 1000              &           &                 &        &     
\end{tabular}
\end{table*}
